---
title: "Transboundary Stocks"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(warning=FALSE, message=FALSE)

library(data.table)
library(rgdal)
library(ggplot2)
library(RColorBrewer)
library(raster)
library(dplyr)

cols      = rev(colorRampPalette(brewer.pal(11, 'Spectral'))(255)) # rainbow color scheme

dir_M = c('Windows' = '//mazu.nceas.ucsb.edu/ohi',
          'Darwin'  = '/Volumes/ohi',
          'Linux'   = '/home/shares/ohi')[[ Sys.info()[['sysname']] ]]

```

### Get species list

```{r sp list}

spp = read.csv('UpsideSpeciesList.csv')%>%
        `colnames<-`(c('x','species'))%>%
          dplyr::select(species)

```

### Grab AquaMaps data

```{r read in aquamaps}

am_sp = read.csv('aquamaps/speciesoccursum.csv')%>%
          mutate(fullname = paste0(genus," ",species))%>%
            filter(fullname %in% spp$species)
am_cells = read.csv('aquamaps/hcaf_truncated.csv')

# am_sp_cells = fread(file.path(dir_M, 'git-annex/globalprep/_raw_data/aquamaps/d2015/csv/hcaf_sp_native_trunc.csv'),
#                     stringsAsFactors=F)%>%
#                   filter(speciesid %in% am_sp$speciesid)%>%
#                     write.csv(file='aquamaps/am_spp_cells.csv')
# 

am_sp_cells = read.csv('aquamaps/am_spp_cells.csv')

```

After filtering the aquamaps species, it looks like `r nrow(am_sp)` of the `r nrow(spp)` RAM species are also in the AquaMaps database.

```{r aquamaps raster}

cells_r = raster('spatial/loiczid_raster.gri')

```


### Rasterize EEZs (OHI regions) to the half degree cells, same as aquamaps. With rgn_id as zones

```{r ohi rgns}

#eezs_r = raster('spatial/rgn_offshore_gcs.tif')%>%resample(.,cells_r,method='ngb',filename='eezs.tif')
eezs_r  = raster('eezs.tif')
plot(eezs_r,col=cols,main='OHI Regions',box=F,axes=F)

```

### Rasterize each aquamap species, and then run zonal on each and get total number of eezs each species has 

```{r}

### Species Map Function ###
# This function takes a single species scientific name as input, then grabs all occurrence cells and associated probability per cell

sp_map_fun <- function(species){
  
  sp = species
  
  sp_id  = filter(am_sp,fullname==sp)$speciesid
  sp_map = filter(am_sp_cells,as.character(speciesid)==as.character(sp_id))%>%
    merge(am_cells,by='loiczid')
  
  return(sp_map)
}

```


### For loop to get EEZs per species

For each species in am_sp and each probability:  

**(1)** set threshold (probability = 0 or 0.5) 
**(2)** return all EEZ rgn_ids  


```{r,eval=F}

probs = c(0,0.5)

#initialize empty dataframe
df = data.frame()

for (i in 1:nrow(am_sp)){
    
  print(i)
  
  #get species name
  sp = am_sp[i,]$fullname
  
  sp_prob_df = data.frame()
  
  for(j in probs){

  
  #get the species map and set as data frame
  
  #I added a column 'presence' set to 1 and removed probability. For some species, like Thunnus obesus, there were duplicate
  # cells with different probabilities which is a problem for the subs() function below... not sure why there are duplicate
  # cells with different probabilities so this is one way around it. When we care about probability of occurrence, we will need
  # to find a better way (maybe take the average prob?)
   map  = sp_map_fun(sp)%>%as.data.frame()%>%
          dplyr::select(loiczid,probability)%>%
            group_by(loiczid)%>%
              summarise(probability = min(probability))%>%
              as.data.frame()%>%
            filter(probability>=j)

  #substitute the species range map cells into the loiczid raster 
  r_sp = subs(cells_r, map[,c('loiczid','probability')], by='loiczid', which='probability', subsWithNA=T)%>%
            mask(eezs_r) #masking the species map with the eez raster in order to remove those cells that are in the species range
                         # map but not included as 'ocean' cells in the eez map. This is not an ideal solution but will 
                         #give a better proportion
  
  #mask OHI EEZ regions with the species map and get unique values 
  m = mask(eezs_r,r_sp)%>%
          unique()
              
  
  #create the data we want - calculate the proportion of cells in each zone.
  d = as.data.frame(m)%>%
        rename(ohi_rgn = m)%>%
        mutate(species = sp)%>%
        left_join(rgns)%>%
        mutate(probability = j)
  
  sp_prob_df = rbind(d,sp_prob_df)
  
  }
  
  df = rbind(sp_prob_df,df)
  
}

#write.csv(df,file='intData/sp_probs_eez_int.csv')

```

### Add Country names to data frame

```{r}

df = read.csv('intData/sp_probs_eez_int.csv',stringsAsFactors=F)%>%
        dplyr::select(-X)
fao_areas = read.csv('fao_areas.csv',stringsAsFactors=F)%>%
            dplyr::select(-X)


data = df%>%
        left_join(fao_areas, by = c("rgn_nam" = "name"))%>%
          rename(FAOarea = num)


#write.csv(data,'sp_prop_eezs.csv')
              

```

****

### Exploratory figures

Number of species with multiple EEZs

```{r}

d = data%>%
    group_by(species,probability)%>%
  summarise(numEEZs = n())%>%
      ungroup()%>%
      as.data.frame()

ggplot(d,aes(numEEZs))+
  geom_histogram(binwidth=5, colour="black", fill="lightblue",alpha=0.4)+
  labs(list(x = "Number of EEZs", y = "Number of Species"))
```

```{r}
ctry = data%>%
        filter(probability==0)%>%
        group_by(rgn_nam)%>%
  summarise(numSPP = n())%>%
  ungroup()%>%
    arrange(desc(numSPP))%>%
  top_n(10)

ggplot(ctry,aes(x = reorder(rgn_nam, numSPP), y=numSPP))+
  geom_bar(position="dodge",stat='identity',fill='lightblue',alpha=0.7,colour="black")+
   theme(axis.text.x=element_text(angle=45,hjust=1))+
  labs(list(x = "Countries", y = "Number of Species (using 0 probability threshold)"))+
  coord_flip()

```




